Renewal and stability in populations structured by remaining years of life

### 2014-09-11 ThanoRepro/ThanoRepro.tex
added caveat to r proof

\section{Equivalence in stability}
\label{app:B}
This appendix contains a proof that the thanatological renewal equation implies
the same intrinsic growth rate, $r$, as the Lotka-Euler model if
the starting population is stable. This is identical to claiming that the right
side of the rewal equation is equal given the same $r$. Since there is a unique
real solution, demonstrated in the previous appendix, then it is sufficient to
show that the two integral equations are equal under these conditions. The basic
relationship is:
\begin{align}
\int _{a=0}^\infty l(a)e^{-ra}f(a) \dd a &= \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra}f^\star (y) \dd a \dd y
\intertext{when}
f^\star(y) &= \frac{b^\star(y)}{c^\star(y)} \\
b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y)\frac{l(a+y)}{l(a)} \dd a \\
c^\star (y) &= \int_{a=0}^\infty c(a) \mu(a+y)\frac{l(a+y)}{l(a)} \dd a
\intertext{Note that:}
c(a) &= \frac{l(a)e^{-ra}}{\int l(a) e^{-ra} \dd a} \\
b(a) &= f(a)c(a)
\end{align}Replacing $f^\star (y)$ with the full expanded
expression and plugging into the thanatological renewal equation gives:
\begin{align}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{\frac{\int_{t=0}^\infty
f(t)l(t)e^{-rt}\mu(a+y)\frac{l(t+y)}{l(t)} \dd t}{\int l(t) e^{-rt} \dd
t}}{\frac{\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{\int
l(t) e^{-rt} \dd t}} \dd a \dd y
\intertext{The denominator of $c(a)$ cancels out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
f(t)l(t)e^{-rt}\mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{
\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t} \dd a \dd y
\intertext{Some $l(a)$'s also cancel out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
f(t)e^{-rt}\mu(t+y)l(t+y) \dd t}{
\int_{t=0}^\infty e^{-rt} \mu(t+y)l(t+y) \dd t} \dd a \dd y
\intertext{Now $d(a+y) = \mu (a+y)l(a+y)$, so we can cancel the inner integral
with the denominator:}
=& \int _{y=0}^\infty 
\int_{a=0}^\infty
f(a)e^{-ra}\mu(a+y)l(a+y) \dd a \dd y
\intertext{Note also that $l(a) = \int _{=a}^\infty d(x) \dd x$ (present
livings are future deaths), which brings us back to the chronological
formulation:} =& \int_{a=0}^\infty e^{-ra}l(a)f(a)
\dd a
\end{align}

>>>>

\section{Equivalence in stability}
\label{app:B}
This appendix contains a proof that the thanatological renewal equation implies
the same intrinsic growth rate, $r$, as the Lotka-Euler model if
the starting population is stable, given a particular set of thanatological
fertility rates.
This is identical to claiming that the right side of the rewal equation is equal given the same $r$. Since there is a unique
real solution, demonstrated in the previous appendix, then it is sufficient to
show that the two integral equations are equal under these conditions. The basic
relationship is:
\begin{align}
\int _{a=0}^\infty l(a)e^{-ra}f(a) \dd a &= \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra}f^\star (y) \dd a \dd y
\intertext{when}
f^\star(y) &= \frac{b^\star(y)}{c^\star(y)} \\
b^\star (y) &= \int_{a=0}^\infty b(a) \mu(a+y)\frac{l(a+y)}{l(a)} \dd a \\
c^\star (y) &= \int_{a=0}^\infty c(a) \mu(a+y)\frac{l(a+y)}{l(a)} \dd a
\intertext{Note that:}
c(a) &= \frac{l(a)e^{-ra}}{\int l(a) e^{-ra} \dd a} \\
b(a) &= f(a)c(a)
\end{align}Replacing $f^\star (y)$ with the full expanded
expression and plugging into the thanatological renewal equation gives:
\begin{align}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{\frac{\int_{t=0}^\infty
f(t)l(t)e^{-rt}\mu(a+y)\frac{l(t+y)}{l(t)} \dd t}{\int l(t) e^{-rt} \dd
t}}{\frac{\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{\int
l(t) e^{-rt} \dd t}} \dd a \dd y
\intertext{The denominator of $c(a)$ cancels out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
f(t)l(t)e^{-rt}\mu(t+y)\frac{l(t+y)}{l(t)} \dd t}{
\int_{t=0}^\infty l(t)e^{-rt} \mu(t+y)\frac{l(t+y)}{l(t)} \dd t} \dd a \dd y
\intertext{Some $l(a)$'s also cancel out:}
=& \int _{y=0}^\infty \int
_{a=0}^\infty d(a+y)e^{-ra} \frac{
\int_{t=0}^\infty
f(t)e^{-rt}\mu(t+y)l(t+y) \dd t}{
\int_{t=0}^\infty e^{-rt} \mu(t+y)l(t+y) \dd t} \dd a \dd y
\intertext{Now $d(a+y) = \mu (a+y)l(a+y)$, so we can cancel the inner integral
with the denominator:}
=& \int _{y=0}^\infty 
\int_{a=0}^\infty
f(a)e^{-ra}\mu(a+y)l(a+y) \dd a \dd y
\intertext{Note also that $l(a) = \int _{=a}^\infty d(x) \dd x$ (present
livings are future deaths), which brings us back to the chronological
formulation:} =& \int_{a=0}^\infty e^{-ra}l(a)f(a)
\dd a
\end{align}
This finding essentially proves that in the stable population there is at least
one thanatological fertility rate schedule that will satisfy the constraints of
the lifespan distribution and $r$. However, it is not necessary to derive the
thanatological fertility distribution in the way prescribed here, and indeed an
infinite number of such fertility distributions would satisfy the same
constraints. Imagine these fertility rates cross-classified in
both age dimensions, where two marginal sums give the thanatological and
chronological rates, as they would be calculated over the
corresponding cross-classified exposures. The shape of this rate surface can be
shifted around in infinitely many ways that would preserve a given growth rate
and lifespan distribution, and the above proof shows one such mapping.
